---
title: "Abalone Report 1"
author: "Melinda Higgins"
date: "2023-08-21"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# set basic options
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(knitr)
library(printr)

# load data
library(readr)
abalone <- read_csv("abalone.csv")
```

## Take a look at the abalone dataset

Using the `head()` function, let's look ta the top 10 rows of the abalone dataset

```{r}
# take a look at top 10 rows of abalone dataset
head(abalone, 10)
```

## Abalone sex frequencies

We'll use the `table()` function to get the counts of the abalone sex

```{r}
# get frequency of sex
table(abalone$sex)
```

## Get some summary stats

Get the mean and standard deviation of abalone lengths

```{r}
# get mean and sd of length
mean(abalone$length)
sd(abalone$length)
```

## Plot of abalone weights and number of rings

Basic plot of abalone whole weights on the x-axis and the number of rings on the y-axis. We'll add colors for the sex of the abalones.

```{r}
# create colors for each sex
# use ifelse() function
# set M to blue
# set F to red
# set I to green
abalone$sexColor <-
  ifelse(
    abalone$sex == "M",
    "blue",
    ifelse(
      abalone$sex == "F",
      "red",
      "green"
  ))

# make plot of x=wholeWeight, y=rings
# color by sex
plot(x=abalone$wholeWeight, 
     y=abalone$rings, 
     pch=16,
     col = abalone$sexColor)
```

